---
title: "Intro to caladaptr"
author: "Andy Lyons"
date: "12/18/2019"
output: 
  html_document: 
    smart: no
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: console
---
<style>
h1 {margin-top:2em; font-size:2em; font-weight:bold; color:darkgreen;}
h2 {font-size:1.5em; color:darkred;}
p, li {font-size:20px;}
.tryit {margin: 1em 2em; padding:0.5em; border:2px solid darkred; color:darkred; font-size:90%;}
li.tocify-item {font-size:100%;}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Background

## Goals of the <tt>caladaptr</tt> R Package

- <tt>caladaptr</tt> is an **API client package** for R.

    - Primary role is to provide low-level functions for **querying** and **importing** Cal-Adapt data through the API.

    - Secondary roles might be computing generic derived dataset not yet available on the API (e.g., growing degree days).

- The intended audience is **experienced R users** (not necessarily climate scientists).

- Accepts and returns standard data classes (e.g., sf, raster)

- Built using guidelines from [rOpenSci](https://devguide.ropensci.org/){target="_blank"} with plans to submit the package for peer review. 

- Enable light-weight decision support tools through RShiny that present Cal-Adapt data in user friendly terminology, combine with other datasets, produce user friendly output, etc. 

- Plans for a companion ebook with code recipes (created with bookdown, [example](https://ucanr-igis.github.io/360photos/){target="_blank"})

- Earlier cookbook: https://ucanr-igis.github.io/caladaptr-dev/cal-adapt-api_cookbook.html

# Installation

Because the package is currently in a private GitHub repo, you need to 

1. Get yourself added to the repo as an editor
2. Generate for yourself a Personal Access Token (with the repo scope  enabled)

```{r get_github_token, eval=FALSE}
gh_token_fn <- "~/My Keys/github_pat.txt"  ## For safety the token is saved in the first line of a text filee
f <- file(gh_token_fn, open = "rt")           
andys_github_pat <- readLines(f, n=1)      ## n=1 read first line only
close(f)
# remotes::install_github("UCANR-IGIS/caladaptr", auth_token = andys_github_pat)
```

Load the package:

```{r install_pkg, cache=TRUE, eval=TRUE}
library(caladaptr)
packageVersion("caladaptr")
```

# Summary of Features

Only working with raster series for now

Only using querying for now (time series end point)

 - downloading rasters will come later

Users can select which raster series they want to get by passinig by the component pieces (GCM + scenario + climate variable + temporal aggregation)

 - built-in constants so they don't have to remember spellings

Querying

   - [`units`](https://rdrr.io/cran/units/) package used to record units (e.g., Kelvin)

   - For now just points are supported  
       - sf package  
       - future: zip codes, counties, Watersheds, congressional districts

# Constants

```{r shohwconstants, cache=TRUE, eval=TRUE}
ca_climvar
ca_gcm
ca_scenario
ca_period
```

# Functions

`ca_rs_getinfo`  
`ca_rs_url`
`ca_fmtdate`
`ca_res_getinfo`  

## Function Naming Logic

- All constants and functions start with <em>ca_</em>

- Second piece specifies dataset(s):

    - <em>_rs</em> raster series  

- Third part is the verb:

    - <em>_getinfo</em> retrieve metadata  

    - <em>_getvals</em> retrieve values  

    - <em>_getrst</em> retrieve raster  


# `ca_rs_getinfo`

Returns a data frame containing info about all raster series

Cached version saved in package

```{r}
args(caladaptr::ca_rs_getinfo)

## Refresh
## y <- ca_rs_getinfo(retrieve=TRUE)

## View to help find a raster series
#View(y)
```

# `ca_rs_url`

Returns a URL (slug) for a raster series given CGM, RCP, etc.

Cross-chcecks against available raster series

Used as a helper function

```{r ca_rs_url, cache=TRUE}
args(caladaptr::ca_rs_url)

## Compute some URLs
ca_rs_url(climvar=c("tasmax", "tasmin"), scenario = ca_scenario[1:2], period="year")

## What happens if you try to get raster series that don't exist??
ca_rs_url(climvar=c("tasmax", "tasmin"), scenario = ca_scenario[1:2], period="month")

```

# `ca_rs_getvals`

Returns time series values for a location

```{r ca_rs_getvals, cache=TRUE}
args(caladaptr::ca_rs_getvals)
```

`loc` will accept a numeric vector (long, lat), sf object, reference to a pre-built vector layer, etc.

`stat` will be a summary function

## Query a Point Location

```{r}
library(caladaptr)

## Define a Point
pt <- c(-121.4687,38.5938)

## Define rasters
data_urls <- ca_rs_url(climvar = "tasmax", period="day", gcm="HadGEM2-ES", scenario = c("rcp45", "rcp85"))
data_urls

## Define date range
dates <- as.Date(c("2050-01-01", "2050-01-31"))

## Get the values for this point
my_vals <- ca_rs_getvals(data_urls, loc=pt, dates=dates)
names(my_vals)
hist(my_vals[["tasmax_day_HadGEM2-ES_rcp45"]]$val)
hist(my_vals[["tasmax_day_HadGEM2-ES_rcp85"]]$val)

```





# Next Steps

- Expand support for `loc` argument

    - sf Polygons
    - Built-in polygon layers (e.g. Counties, Census Tracts) 
    - Zip codes  
    - Multiple locations


# Questions for the Team

1. Any thoughts / preferences about license? Currently using [CC0](https://creativecommons.org/publicdomain/zero/1.0/)



